A Context Menu for the Real World: Controlling Physical Devices Through Head-Worn Infrared Targeting

### 2013-09-14 doc/sigchi14/device.tex
more text and figures
various improvements

\section{Hardware Device and Interaction}

Our prototype consists of a Google Glass Explorer Edition head-worn computing device, augmented with an infrared emitter that is mounted on the frame, pointing out in the direction of the wearer's view (Figure~\ref{fig:glass}). The IR emitter LED is mounted in an opaque hollow tube, that restricts the outgoing angle of illumination.

@@

\subsection{Prototype Implementation}
In our prototype, Glass communicates over Bluetooth to an additional microcontroller board the user has to wear (Atmel ATMega256). This board marshals XBee to Bluetooth messages in both directions and also controls the IR LED mounted on the Glass frame (Figure~\ref{fig:architecture}). This architecture was mostly chosen for reasons of expediency. We selected XBee 802.15.4 radios to avoid the latency associated with connecting and disconnecting to Bluetooth devices \bjoern{Is this true or not?} \sean{Origainlly we didn't know we'd work on Google Glass, therefore BT didn't have a higher priority. We learned that bluetooth can concurrently communicate with up to 7 devices (while up to 255 can be paired but inactive), giving us more complexity when scaling up. In addition ZigBee has shorter wakeup delay and lower power consumption. (referenced here: http://superuser.com/questions/332767/limit-to-the-number-of-devices-that-can-be-paired-with-a-bluetooth-device)} Future head-mounted devices could clearly integrate IR emitters; the choice of local wireless technology could also change. In particular, one could substitute WiFi modules.

@@

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/glass-ir-coverage}
\caption{Our augmented Glass prototype has a frame-mounted infrared emitter.}
\label{fig:coverage}
\end{figure}

>>>>

\section{Interaction and Hardware Device}

\subsection{Design Goals}
{\bf Leverage visual attention:} Take advantage of the fact that visual attention can express intention - trigger interaction based on where a user is already looking

{\bf Provide feedback in the environment:} While a near-eye display can push information to the user, locate visual feedback about selection targets in the environment, to prevent distraction and interruption.

{\bf Deal with dense device arrangements and the limit of head positioning}

\bjoern{Finish and expand this section.}

@@

\subsection{Prototype Implementation}
Our prototype consists of a Google Glass Explorer Edition head-worn computing device, augmented with an infrared emitter that is mounted on the frame, pointing out in the direction of the wearer's view (Figure~\ref{fig:glass}). The IR emitter LED is mounted in an opaque hollow tube, that restricts the outgoing angle of illumination.

In our prototype, Glass communicates over Bluetooth to an additional microcontroller board the user has to wear (Atmel ATMega256). This board marshals XBee to Bluetooth messages in both directions and also controls the IR LED mounted on the Glass frame (Figure~\ref{fig:architecture}). This architecture was mostly chosen for reasons of expediency. We selected XBee 802.15.4 radios to avoid the latency associated with connecting and disconnecting to Bluetooth devices \bjoern{Is this true or not?} \sean{Origainlly we didn't know we'd work on Google Glass, therefore BT didn't have a higher priority. We learned that bluetooth can concurrently communicate with up to 7 devices (while up to 255 can be paired but inactive), giving us more complexity when scaling up. In addition ZigBee has shorter wakeup delay and lower power consumption. (referenced here: http://superuser.com/questions/332767/limit-to-the-number-of-devices-that-can-be-paired-with-a-bluetooth-device)} Future head-mounted devices could clearly integrate IR emitters; the choice of local wireless technology could also change. In particular, one could substitute WiFi modules.

@@

\begin{figure}[t]
\centering
\includegraphics[width=1.0\columnwidth]{figures/glass-ir-coverage}
\caption{Different IR configurations suggest that usable beam widths of 2 to 4 feet and distances up to 16 feet }
\label{fig:coverage}
\end{figure}

### 2013-09-14 doc/sigchi14/introduction.tex
various improvements

\section{Introduction}
\bjoern{What's the overall motivation for the project? What problem does it address?}
Increasingly, devices and services in our built environment are networked and can be controlled remotely. The proliferation of smart, controllable devices such as intelligent lighting, AV equipment, HVAC systems, or kitchen appliances raises the question of how to best interact with them. Prior work has introduced several approaches that use handheld mobile devices as {\em universal remote controls} to select and control appliances in the proximate environment- e.g., through pointing at such devices to select them, and through touch-screen or button interfaces to send control commands~\cite{beigl_point_1999,patel_2-way_2003}.

\bjoern{what are the drawbacks of this existing work?}
Some drawbacks of using handheld devices are that the device first has to be retrieved (e.g., from a pocket) and aimed; that two hands may be necessary for operation (one to hold the device, one to operate the touch screen); and that the user's visual attention is split between looking down at a screen and out at the device to-be-controlled. \bjoern{these all sound kind of weak.} These general limitations have led researchers to investigate ``always-available'' physiological input modalities for phones on one hand~\cite{morris_emerging_2011}, and wearable computing devices on the other hand.

@@

In this paper, we introduce a novel method for selecting and controlling smart appliances in physical spaces though the use of a head-worn computing device with near-eye display and wireless communication. We augment Google Glass\footnote{\url{http://www.google.com/glass/start/}} with custom hardware for this purpose. Users first look in the direction of the device they wish to control to initiate interaction (e.g., at a lamp to control lighting, or at a speaker to change music playback volume). If multiple devices fall within communication range, an on-screen disambiguation dialog lets users select a single target. Once acquired, a device specific control UI shown on the head-mounted display enables adjustment of discrete and continuous parameters through a touchpad interface (see Figure~\ref{fig:teaser}).

Our hardware relies on infrared communication between Glass and target devices to establish a connection; and on wireless ZigBee radio communication to exchange control messages.  Glass is augmented with a narrow IR emitter and a ZigBee ratio. Target devices similarly have IR receivers and ZigBee radios. These choices offer selection-through-head-orientation to initiate interaction; once initiated, users can look away from the target device while issuing control commands.

To understand the system performance, usability, and user experience of head-orientation targeting and head-mounted displays for device control, we first report measurements of the target range and accuracy of our device. We then conduct a comparative study of device acquisition time and error of different interface variants. We find that target acquisition through head orientation is faster than selecting items from a list, given the constraints of linear input using a head-worn touch controller. Finally, we report high-level feedback from $n$ users who use our system for home automation tasks.

>>>>

\section{Introduction}
Increasingly, devices and services in our built environment are networked and can be controlled remotely. The proliferation of smart, controllable devices such as intelligent lighting, AV equipment, HVAC systems, or kitchen appliances raises the question of how to best interact with them.

Today, commercial solutions use handheld mobile devices as {\em universal remote controls} to control such appliances~\bjoern{refs!}. Commonly, users first browse a list of all available devices, and then call up a device-specific user interface. This method faces two challenges: {\em naming} and {\em scoping}. Assigning clear names is non-trivial. In shared spaces, the person trying to control the device might not be the one that named it - e.g., while an office building manager may know what ``Light 4 in area E'' corresponds to, an occupant may not. Second, without a method of scoping selection to automatically filter non-relevant devices, paging though long lists of names or navigating hierarchies becomes potentially more cumbersome than the physical action the ``convenient'' software solution was meant to replace.

To address these challenges, research has introduced techniques of augmenting mobile devices, e.g., with laser pointers, to enable direct aiming at target devices~\cite{beigl_point_1999,patel_2-way_2003}. While promising, some drawbacks of using handheld devices are that the device first has to be retrieved (e.g., from a pocket) and aimed; that two hands may be necessary for operation (one to hold the device, one to operate the touch screen); and that the user's visual attention is now split between looking down at a screen and out at the device to-be-controlled.

@@

In this paper, we introduce a novel method for selecting and controlling smart appliances in physical spaces though the use of a head-worn computing device with near-eye display and wireless communication. We augment Google Glass\footnote{\url{http://www.google.com/glass/start/}} with custom hardware for this purpose. Users first look in the direction of the device they wish to control to initiate interaction (e.g., at a lamp to control lighting, or at a speaker to change music playback volume).  If multiple devices fall within communication range, an disambiguation technique that combines on-screen information as well as visual feedback on the target devices lets users select their desired target. Once acquired, a device specific control UI shown on the head-mounted display enables adjustment of discrete and continuous parameters through a touchpad interface (see Figure~\ref{fig:teaser}).

Our hardware relies on infrared communication between Glass and target devices to establish a connection; and on wireless 802.15.4 radio communication to exchange control messages.  Glass is augmented with a narrow-beam IR emitter and a 802.15.4 radio. Target devices similarly have IR receivers and radios. This combination enables users to initiate interaction by orienting their head; but once initiated, users are free to look away from the target device while issuing control commands.

While prior work has tended to focus on proofs-of-concept, we also contribute empirical data on the system performance, usability, and user experience of  head-orientation targeting and device control. We first report measurements of range and beam characteristics of our controller. We then conduct a study that compares acquisition times for physical targets in a room for our technique and an alternative list selection interface. We find that target acquisition through head orientation is preferred by users and is faster than list selection, given the constraints of linear input using a head-worn touch controller. We also report qualitative results from $n$ participants who use our system for home automation tasks.

### 2013-09-14 doc/sigchi14/relatedwork.tex
added some refs
various improvements

\subsection{Remote Control of Physical Appliances}
Universal device control applications for smart phones are commercially available (e.g., Belkin WeMo\footnote{http://www.belkin.com/us/wemo}), though they usually do not offer spatial selection of target devices, forcing users to browse through lists of configured devices instead. Rukzio found that users strongly preferred either touching a mobile device to a target appliance or pointing at a distance to list browsing~\cite{rukzio_experimental_2006}.

@@

In several techniques, users select objects of interest with laser pointers. The laser dot provides immediate visual feedback to the user what is being selected. Beigl's early AIDA handheld combines laser pointing with IR communication to exchange commands~\cite{beigl_point_1999}. Patel extends this technique by modulating the laser light to communicate the controllers' identity~\cite{patel_2-way_2003} to initiate radio communication. These proofs-of-concept do not include thorough evaluations. Kemp et al. use a laser pointer to indicate to robots which item to pick up in a room~\cite{kemp_point-and-click_2008}.

@@

Standard infrared remote controls for televisions and AV equipment are only meant to control a single device and it is unnecessary to have the user point precisely. These controllers tend to use wide-angle infrared LEDs.

Our main area of differentiation is that we employ head orientation as the selection mechanism instead of pointing - the user looks at the target device to initiate interaction. Selection techniques with very small selectors such as laser dots are less appropriate for head-mounted applications as head orientation only indicates a general area of visual interest. It does not necessarily match gaze orientation as extra-ocular muscles can move the eyes. We therefore select a source with a wider angle of illumination (an IR LED), but restrict its angle to be narrower than in general purpose IR applications.

@@

On the other hand, Card et al. experimentally determined that the bandwidth of neck muscles is much lower than that of arm, wrist or finger muscle groups~\cite{card_morphological_1991}, which limits the performance of any head orientation-based interaction scheme. However, many other factors such as device characteristics and device acquisition time (e.g., pulling a phone out of one's pocket) contribute to overall performance and preference of different selection techniques. Compared to a screen where every pixel is a potential target, the required accuracy for physical device selection in a room is much lower, and head orientation may provide sufficient accuracy.

@@

\subsection{Augmented Reality Interfaces}
Agumented reality applications overlay digital information and graphics on the real world, e.g., through head-mounted displays~\cite{azuma_recent_2001} or other wearable devices. 
Our work is somewhat orthogonal to the research focus of this field as our device's graphics are shown in the visual periphery; they are not referenced to particular objects in the world.

%\bjoern{most of the prior head-mounted work is in the AR space. I don't know much about it yet, but Kortuem~\cite{kortuem_context-aware_1998} might be a good starting point.}

>>>>

\subsection{Remote Control of Physical Appliances}
Standard infrared remote controls for televisions and AV equipment are only meant to control a single device. These controllers tend to use wide-angle infrared LEDs. Universal remote controls are available as dedicated devices (e.g, Crestron\footnote{\url{http://www.crestron.com}}) or applications for smart phones (e.g., Belkin WeMo\footnote{http://www.belkin.com/us/wemo}). They do not offer spatial selection of target devices, forcing users to browse through lists of pre-configured devices instead. Rukzio found that users strongly preferred either touching a mobile device to a target appliance or pointing at a distance to list browsing~\cite{rukzio_experimental_2006}.

@@

In several techniques, users select objects of interest with laser pointers. The laser dot provides immediate visual feedback to the user what is being selected (though it does not indicate whether the pointed-at object can indeed be controlled). Beigl's early AIDA handheld combines laser pointing with IR communication to exchange commands~\cite{beigl_point_1999}. Patel extends this technique by modulating the laser light to communicate the controllers' identity~\cite{patel_2-way_2003} to initiate radio communication. These proofs-of-concept do not include thorough evaluations. Kemp et al. use a laser pointer to indicate to robots which item to pick up in a room~\cite{kemp_point-and-click_2008}.

@@

Our main area of differentiation is that we employ head orientation as the selection mechanism instead of pointing --- the user looks at the target device to initiate interaction. Selection techniques with very small selectors such as laser dots are less appropriate for head-mounted applications as head orientation only indicates a general area of visual interest. It does not necessarily match gaze orientation as extra-ocular muscles can move the eyes. We therefore select a source with a wider angle of illumination (an IR LED), but restrict its angle to be narrower than in general purpose IR applications.

@@

On the other hand, Card et al. experimentally determined that the bandwidth of neck muscles is much lower than that of arm, wrist or finger muscle groups~\cite{card_morphological_1991}, which limits the performance of any head orientation-based interaction scheme. However, many other factors such as device characteristics and device acquisition time (e.g., pulling a phone out of one's pocket) contribute to overall performance and preference of different selection techniques.  Compared to a screen where every pixel is a potential target, the required accuracy for physical device selection in a room is much lower, and head orientation may provide sufficient accuracy. Our work only uses head orientation for the initial selection step; since we believe a user's attention is often drawn to the objects they intend to interact with. Further interactions do not use head orientation.

@@

\subsection{Augmented Reality Interfaces}
Agumented reality applications overlay digital information and graphics on the real world, e.g., through head-mounted displays~\cite{azuma_recent_2001} or other wearable devices. 
Our work is somewhat orthogonal to the research focus of this field as our device's graphics are shown in the visual periphery; they are not referenced to particular objects in the world, though our techniques could be extended to such configurations.

\bjoern{add something about the assembly instruction work here.}
%\bjoern{most of the prior head-mounted work is in the AR space. I don't know much about it yet, but Kortuem~\cite{kortuem_context-aware_1998} might be a good starting point.}
